MCML with Boundary Surfaces

### 2019-05-20 docs/report/article.tex
Update report

\begin{abstract}
	This project report describes an extended implementation of multi-layered Monte Carlo photon transport\cite{wang1992monte}, abbreviated as MCML. It supports structured boundaries, whereas original MCML only supported planar layers. This allows to use the accurate Monte Carlo method for simulation of light in more realistic materials. The report presents and compares radiance profiles of different boundary shapes. The focus lies on the properties reflectance and transmittance, that are important for visual appearance.
\end{abstract}

@@

The report will first explain the principles and purpose of Monte Carlo light transport, before moving on to related work covering MCML specifically. On this basis, the contributions made by this project are stated and explained in detail in the implementation section. The new implementation is verified against the existing work and finally results are presented.

\subsection{Monte Carlo Light Transport}

@@

Photons are initialized with a weight of $1.0$, which is explained in \autoref{impl:photon-transport}. Then, the photon performs a series of repeated steps. Each step length is generated by drawing a random number from a probability distribution based on the interaction coefficient $\mu_a+\mu_s$. The photon is moved by this length in the current direction. The moved photon path has to be checked for collision with the layer boundaries. After each step the photon interacts with the material.

@@

An additional data structure is the array of photon states, which is needed to preserve state across GPU runs. A photon state contains current position, direction, layer index, weight and state of the xorshift~\cite{marsaglia2003xorshift} random number generator. When a state is initialized, a hash function is used on the state index to seed the xorshift generator.

@@

# explicit heightfield boundary with 2 samples
b 2 0.2.1 0.0 0.1

>>>>

\begin{abstract}
	This project report describes an extended implementation of multi-layered Monte Carlo photon transport\cite{wang1992monte} (MCML). It supports structured boundaries, whereas original MCML only supported planar layers. This allows to use the method for simulation of light in more realistic materials. The report presents and compares radiance profiles of different boundary shapes. The focus lies on the properties reflectance and transmittance, that are important for visual appearance.
\end{abstract}

@@

The purpose of the MCML implementation presented here, is to provide a ground truth of surface appearance for physically based rendering. Monte Carlo methods are accepted to be leading in accuracy, but low-noise results have to be bought with long computation times. The implementation produces radiance profiles, that reveal the effect of varying surface shapes, and therefore allow to model surface appearance with approximate analytic curves. The curve equations, when found, can be used in faster rendering algorithms.

The report will first explain the principles of Monte Carlo light transport, before moving on to related work covering MCML specifically. On this basis, the contributions made by this project are stated and explained in detail in the implementation section. The new implementation is verified against the existing work and finally results are presented.

\section{Background}

@@

Photons are initialized with a weight of $1.0$. The presence of weight means, that not actually single photons, but photon packets are simulated. A photon packet does not have atomic behavior, in the sense that its energy content can be divided into fractions. However the packet moves on a single path as if it was a single photon. This is done to reduce variance and hereby the overall number of photons needed for low-noise results. In the following, photon and photon packet refer to the same thing.

The photon performs a series of repeated steps. Each step length is generated by drawing a random number from a probability distribution based on the interaction coefficient $\mu_a+\mu_s$. The photon is moved by this length in the current direction. The moved photon path has to be checked for collision with the layer boundaries. After each step the photon interacts with the material.

@@

An additional data structure is the array of photon states, which is needed to preserve state across GPU runs. A photon state contains current position, direction, layer index, weight and state of the random number generator.

A xorshift~\cite{marsaglia2003xorshift} random number generator is used. It takes more processor instructions, but produces higher quality pseudo random numbers than the linear congruential generator, which was used in the original MCML.

When a photon state is initialized, a hash function is used on the state array index to seed the xorshift generator. This ensures well distributed seeds and hereby largely independent random sequences. Care has to be taken, to not seed xorshift with zero, since the sequence would stay zero forever.

@@

# explicit heightfield boundary with 2 samples
b 2 0.2 0.1 0.0 0.1
